---
title: "Data Science Midterm Progress Report"
author: "Sam Goss"
date: "September 22, 2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```



```{r packages}
library( tidyverse )
library( ggplot2 )
eng_rems <- read_csv("engineRemovals20172020.csv", col_names = TRUE)
```

## Premature Engine Removals

### DATA
+ I:\"engineRemovals20172020.csv"
+ Includes data on `r nrow( eng_rems )` engine removals from Jan 2017 - Sep 2020
+ CSV file of excel spreadsheet of removals maintained by Engine Planning Department


+ Sample of Data

```{r samp , include=TRUE}
head(eng_rems)
```

### Initial Questions
+ Are premature engine removal causes random or specific
+ How many flight cycles define a premature removal

### New / Altered Questions
+ What type of removals should be included:
    - LLP - (life limited parts)
    - UER - (unscheduled engine removals)
    - Planned - problem identified, given additional cycles to removal
    - All others were for company convenience and should not be considered.
    - Filter type of removals
```{r echo=TRUE, include=TRUE}
UPLREM <- eng_rems %>%
  filter(`Type of Removal` == "LLP" | `Type of Removal` == "UER" | `Type of Removal` == "Planned")
```
+ How to handle 19 fleet types.
    - Six fleets eliminated due to observations less than 14 (retired or new fleets)
    - Group by aircraft type or OEM
        - CF6 (CF6-80C2) - GE
        - CFM (CFM56) - GE & Safran
        - PW (PW2000 & PW400) - Pratt & Whitney
        - T-Tails (BR715, V2500, JT8D) - largest observation, sunsetting fleets

### Evaluate CF6 Fleet first       
```{r include=TRUE}
CF680 <- UPLREM %>%
  filter(`Engine Type` == "CF6-80C2B6F" | `Engine Type` == "CF6-80C2B6" | `Engine Type` == "CF6-80C2B8F")

```
   
+ Includes `r nrow( CF680 )` removals
```{r include=TRUE}
ggplot(data=CF680) +
  geom_bar(mapping = aes(x=`Engine Type`, fill = `Type of Removal`))
```
```{r include=TRUE}
ggplot(data=CF680, mapping = aes(x=CSR, fill = `Type of Removal`)) +
  geom_histogram(binwidth = 250)
```
    
+ Histogram shows first 500 cycles appear to be outliers - confirm with boxplot
```{r include=TRUE}
ggplot(data = CF680, mapping = aes(x=`Engine Type`, y= CSR)) +
  geom_boxplot()
```

+ Reduce data set to only removals in first 500 cycles
```{r include=TRUE}
CF680LT <- CF680 %>% filter(CSR <= 500)
```
+ Down to `r nrow( CF680LT )` removals
+ What are the causes for these removals?
```{r include=TRUE}
ggplot(data = CF680LT, aes(x=`Engine Type`, fill = Reason)) +
  geom_bar(position = "dodge")
```

## Conclusion 
+ Since no reason has more than 2 removals, the early removals are random.


    